{Hierarchy Builder:}\\ {Algebraic hierarchies} \\ {made easy} {in {\sc Coq} with {\sc Elpi}}

### 2020-07-01 talk-fscd/main.tex
add Date/Time for the Coq WS
last minute fixes

\begin{frame}[fragile]
  \frametitle{Implementations in DTT (unbundled classes)}

@@

\begin{frame}[fragile]
  \frametitle{Implementations in DTT (packed classes)}

@@

  % more details comming
  \begin{itemize}
  \item {\sc Coq/Mathcomp}:     Packed classes\only<3>{\alert{ inside canonical
    structures}}.
\item {\sc Coq/Math-Classes}: Fully unbundled \alt<3>{\alert{type classes}}{records} \\
    \hfill ($+$ special case for varieties).
  \item {\sc Lean3/Mathlib}:     Semi-bundled \alt<3>{\alert{type classes}}{records}.
  \item {\sc Agda}:             Bundled and semi-bundled records.
  \end{itemize}
  \vfill
  \pause

@@

\center{\textbf{Representations work hand in hand with an inference
  mechanism.}}

@@

  \begin{itemize}
  \item {\sc Coq/Mathcomp}:     canonicals \alert<1>{$+$ heavy boilerplate $+$ checker}
  \item {\sc Coq/Math-Classes}: type classes \alert<1>{$+$ boilerplate $+$ hints}
  \item {\sc Lean3/Mathlib}:     type classes \alert<1>{$+$ priorities $+$ linter}
  \item {\sc Agda}:             records \alert<1>{$+$ open and renaming directives}
  \end{itemize}

@@

  \item Euclidean Spaces $\rightarrow$ Normed Spaces $\rightarrow$
    Complete Space $\rightarrow$ Metric Spaces $\rightarrow$ Topological
    Spaces $\rightarrow$ ...

@@

\begin{frame}
  \frametitle{More examples}

@@

\begin{frame}
  \frametitle{Structure extension vs Structure entailment}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{center}
        {\bf \emph{Structure extension}}
      \end{center}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{center}
        {\bf \emph{Structure entailment}}
      \end{center}
    \end{column}
  \end{columns}
  \vfill
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item \textbf{Compositional}: no need to start from scratch every time.
        (E.g. the product of two groups is a group)
        \pause\vfill
      \item \textbf{Noisy}: changes the internal definition of a structure.
        (E.g. defining a commutative monoid from a monoid, one gets an unnecessary axiom),
        \pause\vfill
      \item \textbf{Non-robust}: breakage of user code when adding new
        intermediate structures, \pause\vfill
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item \textbf{Flexible}: no need to cut structures
        into small bits,
        \pause\vfill
      \item \textbf{Robust}: lets us fix the operations and axioms
        once and for all.
        \pause\vfill
      \item \textbf{Not suitable for inference}: Major breakage when
        arbitrary entailment is automatic. (cf IJCAR \emph{Competing
          Inheritance Paths in Dependent Type Theory)} \pause\vfill
      \end{itemize}
    \end{column}
  \end{columns}
  \vfill

@@

  See a bigger demo during the Coq workshop.
\end{frame}

@@

  \textbf{{\sc Elpi} is a programming language,}
  \begin{itemize}
  \item prolog-like: \textbf{programs and data are clauses}
  \item with binders, unification and constraints
    \pause
  \item capable if representing {\sc Coq} terms in HOAS, its typing
    judgements, evaluation and unification.
  \end{itemize}
    \pause
    \vfill

@@

  \vfill\pause
  Also, \textsc{Coq-ELPI} turned out to be a very comfortable
  meta-programming language for this.

@@

\begin{frame}
  \frametitle{Future work on Hierarchy Builder}
  \begin{itemize}
  \item Adding support for parameters. (debugging phase in progress)
    \vfill
  \item Detecting and fixing competing inheritance paths.
    \vfill
  \item Generating hierarchies of morphisms from structures.
    \vfill
  \item Generating hierarchies of subobjects from structures.
    \vfill
  \item Supporting multiple instances on the same carrier.
    \vfill
  \item Replacing all uses in math-comp and extensions. (just started)
    \vfill
  \item Get better error messages. (work in progress)
  \end{itemize}
\end{frame}

>>>>

\begin{frame}[fragile]
  \frametitle{Implementations in DTT (unbundled classes) \\ \hfill [MSCS2011]}

@@

\begin{frame}[fragile]
  \frametitle{Implementations in DTT (packed classes) \\ \hfill [TPHOL
    2009]}

@@

  % more details comming
  \begin{itemize}
  \item {\sc Coq/Mathcomp}:     Packed classes\only<3>{\alert{ inside canonical
    structures}}.
\item {\sc Coq/Math-Classes}: Fully unbundled \alt<3>{\alert{type classes}}{records} \\
    \hfill ($+$ special case for varieties).
  \item {\sc Lean3/Mathlib}:     Semi-bundled \alt<3>{\alert{type classes}}{records}.
  \item {\sc Agda}:             Bundled and semi-bundled records.
  \item ...
  \end{itemize}
  \vfill
  \pause

@@

\center{\textbf{Representations work hand in hand with tooling.}}

@@

  \begin{itemize}
  \item {\sc Coq/Mathcomp}:     canonicals \alert<1>{$+$ boilerplate
      $+$ validator [IJCAR]}
  \item {\sc Coq/Math-Classes}: type classes \alert<1>{$+$ boilerplate $+$ hints}
  \item {\sc Lean3/Mathlib}:     type classes \alert<1>{$+$ priorities $+$ linter}
  \item {\sc Agda}:             records \alert<1>{$+$ open and renaming directives}
  \end{itemize}

@@

  \item Normed Space $\rightarrow$ Metric Spaces $\rightarrow$
    Topological Spaces $\rightarrow$ ...

@@

\begin{frame}
  \frametitle{More examples \hfill [IJCAR]}

@@

\begin{frame}
  \frametitle{Structure extension vs Structure entailment}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{center}
        {\bf \emph{Structure extension}}
      \end{center}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{center}
        {\bf \emph{Structure entailment}}
      \end{center}
    \end{column}
  \end{columns}
  \vfill
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item \textbf{Compositional}: no need to start from scratch every time.
        (E.g. the product of two groups is a group)
        \pause\vfill
      \item \textbf{Noisy} internal definition of a structure.
        (E.g. defining a commutative monoid from a monoid, one gets an unnecessary axiom),
        \pause\vfill
      \item \textbf{Non-robust} when adding new
        intermediate structures, \pause\vfill
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item \textbf{Flexible}: no need to cut structures
        into small bits,
        \pause\vfill
      \item \textbf{Robust}: we can fix operations and axioms
        once and for all.
        \pause\vfill
      \item \textbf{Not suitable for inference}: Major breakage when
        arbitrary entailment is automatic. (cf IJCAR \emph{Competing
          Inheritance Paths in Dependent Type Theory)} \pause\vfill
      \end{itemize}
    \end{column}
  \end{columns}
  \vfill

@@

  A bigger demo at the Coq Workshop, Monday 10AM GMT+2.
\end{frame}

@@

  \textbf{{\sc Elpi} is a programming language,} \hfill [LPAR-20]
  \begin{itemize}
  \item prolog-like: \textbf{programs and data are clauses}
  \item with binders, unification and constraints
    \pause
  \item capable if representing {\sc Coq} terms in HOAS, its typing
    judgements, evaluation and unification.
  \end{itemize}
    \pause
    \vfill

@@

  \vfill\pause
  Also, \textsc{Coq-ELPI} turned out to be a very comfortable
  meta-programming language for this (approx. 2000 loc).

@@

\begin{frame}
  \frametitle{Future work on Hierarchy Builder}
  \begin{itemize}
  \item Adding support for parameters. (debugging phase in progress)
    \vfill
  \item Replacing all uses in math-comp and extensions. (just started)
    \vfill
  \item Get better error messages. (work in progress)
    \vfill
  \item Detecting and fixing competing inheritance paths.
    \vfill
  \item Generating hierarchies of morphisms from structures.
    \vfill
  \item Generating hierarchies of subobjects from structures.
    \vfill
  \item Supporting multiple instances on the same carrier.
  \end{itemize}
\end{frame}
